A minimal reaction-diffusion neural model generates {\emph{C. elegans}} undulation

### 2020-12-25 paper.tex
Some minor changes to phrasing

@@

In the terms of dynamical systems analysis, the oscillations destabilize to a stable limit cycle.  The mutual inhibition and symmetry between neuron pairs leads to a phase lock of $180^\circ$.  This provides out-of-phase stimuli which propagate through the descending pathway, with the inhibition between neuron pairs keeping the oscillations stable.
% TODO experiment on a single head oscillator.
% TODO explain how coupling works

@@

% TODO do we need this?
% One can think of the CPG as a driven, dissipative system.  An analogy to the driven, damped oscillator is particularly apt - in the CPG, the head oscillator drives the descending pathway.

% TODO expand.  Our model could have had more neurons, with a shorter phase lag between pairs.
While \textit{C. elegans} has twelve pairs of motor neurons, we have only used six pairs in our model.  \Fref{fig: cpg} is a depiction of our simplified model as a graph, wherein neurons are nodes, and the arrows between them symbolize connections.

@@

% TODO: remove Morris-Lecar
% remark that if the w and v nullclines come near each other, you see a critical slowdown.
% In Hodgkin-Huxley notation, $n$ is the potassium channel.  In the FHN formulation, this corresponds to the $w$ gate variable.
\begin{figure}[h!]
    \label{fig: nm nullclines}
    \centering
    \includegraphics{figures/neuron_dynamics/neuron_dynamics.pdf}
    % note that in the vicinity of the v-nullcline, the w-nullcline is essentially straight.
    \caption{Nullclines of several different neuronal models; on the right is the biological model of Xu et al, in the centre is the Morris-Lecar model, and to the right is the FitzHugh-Nagumo model.  These have been arranged in order of decreasing complexity.}
\end{figure}

@@

@@

The worm generates sinusoidal oscillations

@@

% TODO: write up a section on the Beer paper and see how it relates to our CPG model.
% Perhaps a discussion about the "scientific lineage" of our model, and the Olivares et al model.
% Paraphrase their description.

@@

Mention the bullet points from the presentation.  How could a system like this be used in applications?  End with some future paths.

>>>>

There are many other models for the CPG; for example, a 2019 preprint by \cite{olivares2019} proposes a distributed network of oscillating systems, instead of a structured chain like our proposed CPG.  The point of this paper is to create a minimal working model, though, and not to strive for full realism.

@@

In the terms of dynamical systems analysis, the oscillations destabilize to a stable limit cycle.  The mutual inhibition and symmetry between neuron pairs leads to a phase lock of $180^\circ$.  This provides out-of-phase stimuli which propagate through the descending pathway, with the inhibition between neuron pairs keeping the oscillations stable.

@@

% TODO expand.  Our model could have had more neurons, with a shorter phase lag between pairs.
While \textit{C. elegans} has twelve pairs of motor neurons, we have only used six pairs in our model.  An equivalent result could be obtained with more neurons, by increasing the coefficient of diffusion along the pathway, causing a shorter phase lag between neuron pairs.  \Fref{fig: cpg} is a depiction of our simplified model as a graph, wherein neurons are nodes, and the arrows between them symbolize connections.

@@

% remark that if the w and v nullclines come near each other, you see a critical slowdown.
% In Hodgkin-Huxley notation, $n$ is the potassium channel.  In the FHN formulation, this corresponds to the $w$ gate variable.
\begin{figure}[h!]
    \label{fig: nm nullclines}
    \centering
    \includegraphics{figures/neuron_dynamics/neuron_dynamics.pdf}
    % note that in the vicinity of the v-nullcline, the w-nullcline is essentially straight.
    \caption{Nullclines of several different neuronal models; on the right is the biological model of Xu et al, in the centre is the Morris-Lecar model, and to the right is the FitzHugh-Nagumo model.  These have been arranged in order of decreasing complexity.}
\end{figure}

@@

\subsection{Generating undulation from neuron potential}

The CPG network which we created is of the same form as the 2-D musculature of our simulated worm.  Each neuron is attached to a muscle; when it fires, it causes the attached muscle to contract.​  We model the muscles as leaky integrators of the neuron's voltage; therefore, Gaussian smoothing of the neuron's voltage works to ``create'' an effective muscle contraction signal.  Through Gaussian smoothing, the movement is heavily dependent on the current state of the neuron but also dependent somewhat on the history. ​

Much like \citet{izquierdo2018} did, we take a segment-based approach to the worm model.  Pairs of muscles on either side determine the overall angular displacement per segment.​  The magnitude of the “contraction” resulting from the smoothing is interpreted as an angular displacement for that segment.​  We use a cubic spline interpolation to smoothen the worm body.

Each neuron is assumed to be connected to a muscle; when it fires, it should cause that muscle to contract.​
Pairs of neurons on either side determine the overall angular displacement per segment.​
Gaussian smoothing and a cubic spline are used to mimic muscle movement and medium resistance.​
The magnitude of the “contraction” resulting from the smoothing is interpreted as an angular displacement for that segment.​

In this model, muscles can be seen as “leaky integrators” (Boyle et al. 2012) – the movement is heavily dependent on the current state of the neuron but is also dependent somewhat on the history.

@@

The worm undulates in a sinusoidal manner, both in simulation and reality.

@@

\subsection{}
% TODO: write up a section on the Beer paper and see how it relates to our CPG model.
% Perhaps a discussion about the "scientific lineage" of our model, and the Olivares et al model.
% Paraphrase their description.

@@
